A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-06 latex/main.tex
Updates from Overleaf

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
Although uncommon relative to single nucleotide mutations,
deletions are an important component of SARS-CoV-2 evolution,
contributing to the emergence of VOCs by enhancing,
e.g., efficiency for cell entry \cite{Cantoni2022} and
potential for antibody escape \cite{McCarthy2021}.
Using the ARG, we identified major deletion events
(which are propagated to many descendants)
as well as recurrent deletions
(which occur on at least two branches) in the ARG.
In Table~\ref{tab:major_dels},
we list the major deletion events in the ARG
which are each inherited by at least 10,000 descendant samples; and
in Table~\ref{tab:rec_dels},
we list the recurrent deletions in the ARG
which are each observed by at least 10,000 (0.4\%) samples.
[FIXME define how we measured this].

@@

\begin{table}
\centering
\caption{
Major deletion events in the ARG.
The deletions caused by these events
are inherited by at least 10,000 samples.
These deletions are characterized by
a site position in the genome (start) and
the number of subsequence sites involved (length), and
they are annotated by the gene in which they occur
(non-coding otherwise).
The IDs of the nodes in the ARG
above which the deletions occur and
the Pango lineages assigned to these nodes are indicated.
The percentage of samples inheriting each deletion
is shown as frequencies.
If a deletion has been identified
as a characteristic mutation of a VOC by a study (or an analysis),
a reference to the study (or the analysis) is provided.
Abbreviations: NSP, non-structural protein;
ORF, open reading frame;
S, Spike;
S1-NTD, N-terminal domain in Spike S1 subunit.
}
\label{tab:major_dels}
\begin{tabular}{rccrcrrc}
\toprule
Start & Length & Region & Node & Pango lineage & Frequency (\%) & Ref\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 45.06 &  \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 44.91 & \cite{Stern2021} \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 44.82 & \cite{Stern2021} \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 (Omicron) & 35.12 &  \\
28362 & 9 & N & 1436802 & B.1.1.529 (Omicron) & 35.12 & \cite{RambautOmicron2021} \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 21.44 & \cite{RambautOmicron2021} \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 21.43 &  \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 13.73 & \cite{RambautOmicron2021} \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 13.69 &  \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 13.68 & \cite{RambautOmicron2021} \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 13.68 & \cite{RambautOmicron2021} \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 12.60 & \cite{RambautOmicron2021} \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 11.70 &  \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 11.70 & \cite{RambautAlpha2020} \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 11.64 & \cite{RambautAlpha2020} \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 11.70 & \cite{RambautAlpha2020} \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 7.67 & \cite{Tegally2022} \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 0.94 &  \\
\bottomrule
\end{tabular}
\end{table}

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
Although uncommon relative to single nucleotide mutations,
deletions are an important component of SARS-CoV-2 evolution,
contributing to the emergence of VOCs by enhancing,
e.g., efficiency for cell entry \cite{Cantoni2022} and
potential for antibody escape \cite{McCarthy2021}.
Using the ARG, we identified major deletion events
(which are propagated to many descendants)
as well as recurrent deletions
(which occur on at least two branches) in the ARG.
In Table~\ref{tab:major_dels},
we list the major deletion events in the ARG
which are each inherited by at least 10,000 descendant samples; and
in Table~\ref{tab:rec_dels},
we list the recurrent deletions in the ARG
which are each observed in at least 10,000 (0.4\%) samples
% [FIXME define how we measured this].
(i.e., for a deletion caused by multiple events in the ARG,
the sum of all the samples inheriting the deletion
across all the events).

@@

\begin{table}
\centering
\caption{
Major deletion events in the ARG.
The deletions caused by these events
are inherited by at least 10,000 samples.
The deletions are characterized by
a site position in the genome (start) and
the number of sites involved (length);
they are annotated by the gene in which they occur
(non-coding otherwise).
The IDs of the nodes in the ARG
above which the deletions occur and
the Pango lineages assigned to these nodes are shown.
Frequency is the percentage of samples inheriting a deletion.
If a deletion has been identified
as a characteristic mutation of a VOC
by a study (or an analysis),
a reference to the study (or the analysis) is provided.
Abbreviations: NSP, non-structural protein;
ORF, open reading frame;
S, Spike;
S1-NTD, N-terminal domain in Spike S1 subunit.
}
\label{tab:major_dels}
\begin{tabular}{rccrcrrc}
\toprule
Start & Length & Region & Node & Pango lineage & Frequency (\%) & Ref\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 45.06 &  \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 44.91 & \cite{Stern2021} \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 44.82 & \cite{Stern2021} \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 (Omicron) & 35.12 &  \\
28362 & 9 & N & 1436802 & B.1.1.529 (Omicron) & 35.12 & \cite{RambautOmicron2021} \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 21.44 & \cite{RambautOmicron2021} \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 21.43 &  \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 13.73 & \cite{RambautOmicron2021} \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 13.69 &  \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 13.68 & \cite{RambautOmicron2021} \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 13.68 & \cite{RambautOmicron2021} \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 12.60 & \cite{RambautOmicron2021} \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 11.70 &  \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 11.70 & \cite{RambautAlpha2020} \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 11.64 & \cite{RambautAlpha2020} \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 11.70 & \cite{RambautAlpha2020} \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 7.67 & \cite{Tegally2022} \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 0.94 &  \\
\bottomrule
\end{tabular}
\end{table}
